Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-18 article.tex
add results

\textcolor{red}{SEUIL 0.11 for tabulars!!!}

In this section, we will first describe datasets and methods used for the comparison. Then, on Tab.\ref{tab:tab3}, our approach is decomposed to show the contribution of each stages . Afterwards, we will perform a comparison between several methods on Tab.\ref{tab:tab1} (or Fig.?). Note that two metrics are used to evaluate the performance. The end of this section is devoted to a discussion about the choice of metric and its relevance for the challenge.

@@

% Metrics : ROC AUC and AUPRC
In supervised learning terms, the network inference task can be viewed as a
binary classification task where one has to correctly classify the presence
or the absence of edges of the graph. We assess the accuracy of the
network inference method using the area under the ROC curve (AUROC)
and the area under the precision-recall curve (AUPRC)
\citep{schrynemackers2013protocols}. Since neurons are not self-connected
($(i, i) \not \in E, \forall i \in V$) and some inference methods could
give maximal scores to the diagonal of $\hat{A}$, e.g. Pearson
correlation, we set those elements to the minimum of $\hat{A}$ prior
assessment.

We compare our approach to the Pearson correlation (see Section
\ref{sec:inference}), the generalized transfer entropy (GTE) and GENIE3.
\begin{description}
\item[GTE], the baseline of the challenge, measures for all pairs of
neurons $i$ and $j$ the reduction of uncertainty in predicting future values of
$i$ given the observation of the time series of the neuron $j$
conditionally to the observation of past time series of neuron $i$,
whenever the average fluorescence is over a conditioning level.
Conditioning level was set to reproduce baseline of the challenge.

\item[GENIE3] (\cite{huynhthu2010inferring}) translates the inference of the
network of $p$ neurons into $p$ supervised learning tasks. For each task,
the fluorescence of a neuron is predicted from the fluorescence
of all other neurons using a tree based ensemble such as
random forest (\cite{breiman2001random}). The tree ensemble associate a
variable importance scores (\cite{louppe2013understanding})
to each neuron used in the predictions, which is used as a connectivity
score. The following parameters are used
$10000$ trees, $\log_2{p}$ neurons are randomly selected at each tree
node to split the data and the maximal tree depth is constrained
to 1, 2 or 3.
\end{description}

@@

At first sight, it appears on Tab.\ref{tab:tab1} that partial
correlation outperforms other methods. Two main reasons may explain this.
First, the quality of the infered network seems related to the conditioning
part of the method. In every method except with partial correlation, indirect
links are confounded with direct links.  Secondly, the gaussian assumption
seems verified and thus linear method appears to be competitive, e.g. the
method correlation already gives good results, with a non-linear method such
as GENIE3.

@@

\begin{table}[tbh]
\centering
\begin{tabular}{@{}l *{8}{c}@{}}
\hline
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
No  filtering       & 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
P                   & 0.925 & 0.925 & 0.924 & 0.923 & 0.322 & 0.324 & 0.323 & 0.312\\
P + W               & 0.931 & 0.930 & 0.928 & 0.927 & 0.333 & 0.334 & 0.327 & 0.313\\
P + W + PCA         & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344\\
Averaging           & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388\\
Averaging + Directivity & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389\\
\end{tabular}
\caption{TODO \textcolor{red}{version simple!!}}
\label{tab:tab3}
\end{table}

\begin{table}
\small
\centering
\begin{tabular}{@{}l *{8}{c}@{}}
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline \hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
\hline \hline
% Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
Correlation            & 0.886 & 0.884 & 0.891 & 0.877 & 0.153 & 0.145 & 0.170 & 0.132 \\
GENIE3 (depth = 1)     & 0.878 & & & & 0.184 & & & \\
GENIE3 (depth = 2)     & 0.889 & & & & 0.217 & & & \\
GENIE3 (depth = 3)     & & & & & & & & \\
GTE                    & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142 \\
Partial correlation    & 0.930 &  0.927 &  0.926 &  0.923& 0.363  & 0.347 &  0.346 & 0.328 \\
Partial correlation (avg.) & 0.938 & & & 0.932& 0.391 & & & 0.374\\
% Our approach simple + dir & 0.939 & & & 0.933& 0.392 & & & 0.375\\
Partial correlation (opt.) & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344 \\
Our approach (opt. \& avg.)    & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388 \\
% Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
\end{tabular}
\caption{TODO}
\label{tab:tab1}
\end{table}

@@

In previous study \citep{shipley2002cause}, it has been suggested to
consider correlation coefficients of all possible orders,
i.e. conditioned on every subsets of the set of $p-2$ other variables. This
would take into account multiple indirect paths between a given pair of
variables.

@@

>>>>

%\textcolor{red}{SEUIL 0.11 for tabulars!!!}

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{images/curves}
\caption{Method precision-recall and roc curves.}
\label{fig:curves}
\end{figure}

In this section, we will first describe datasets and methods used for the
comparison. Then, on Tab.\ref{tab:tab3}, our approach is decomposed to show
the contribution of each stages . Afterwards, we will perform a comparison
between several methods on Fig.\ref{fig:curves} and Tab.\ref{tab:tab1}. Note that two
metrics are used to evaluate the performance and a single low-pass filter is used (Eq.\ref{eq:symetric-median}) and a threshold, i.e. $\tau = 0.11$. The end of this section is
devoted to a discussion about the choice of metric and its relevance for the
challenge.\\

@@

% Metrics : ROC AUC and AUPRC
In supervised learning terms, the network inference task can be viewed as a
binary classification task where one has to correctly classify the presence
or the absence of edges of the graph. We assess the accuracy of the
network inference method using the area under the ROC curve (AUROC)
and the area under the precision-recall curve (AUPRC)
\citep{schrynemackers2013protocols}. Since neurons are not self-connected
($(i, i) \not \in E, \forall i \in V$) and some inference methods could
give maximal scores to the diagonal of $\hat{A}$, e.g. Pearson
correlation, we set those elements to the minimum of $\hat{A}$ prior
assessment.\\

On Tab.\ref{tab:tab3}, it can be seen that each step of our approach globally improves the score on both metrics. However, processing, weights and averaging (respectively denoted as $P$,$W$ and $Averaging$ in Tab.\ref{tab:tab3}) as described in section \ref{sec:filter} are the key steps of our algorithm.

We compare our approach to the Pearson correlation (see Section
\ref{sec:inference}), the generalized transfer entropy (GTE) and GENIE3.

\textbf{GTE} \citep{stetter2012model}
, the baseline of the challenge, measures for all pairs of
neurons $i$ and $j$ the reduction of uncertainty in predicting future values of
$i$ given the observation of the time series of the neuron $j$
conditionally to the observation of past time series of neuron $i$,
whenever the average fluorescence is over a conditioning level.
Conditioning level was set to reproduce baseline of the challenge.

\textbf{GENIE3} \citep{huynhthu2010inferring} translates the inference of the
network of $p$ neurons into $p$ supervised learning tasks. For each task,
the fluorescence of a neuron is predicted from the fluorescence
of all other neurons using a tree based ensemble such as
random forest \citep{breiman2001random}. The tree ensemble associate a
variable importance scores \citep{louppe2013understanding}
to each neuron used in the predictions, which is used as a connectivity
score. The following parameters are used
$10000$ trees, $\log_2{p}$ neurons are randomly selected at each tree
node to split the data and the maximal tree depth is constrained
to 1, 2 or 3.\\

@@

At first sight, it appears on Fig.\ref{fig:curves} %Tab.\ref{tab:tab1}   
that partial correlation outperforms other methods. Two main reasons may explain
this. First, the quality of the infered network seems related to the
conditioning part of the method. In every method except with partial
correlation, indirect links are confounded with direct links.  Secondly, the
gaussian assumption seems verified and thus linear method appears to be
competitive, e.g. the method correlation already gives good results, with a
non-linear method such as GENIE3.

@@

\begin{table}[tbh]
\centering
\begin{tabular}{@{}l *{8}{c}@{}}
\hline
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
No  filtering       & 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
P                   & 0.925 & 0.925 & 0.924 & 0.923 & 0.322 & 0.324 & 0.323 & 0.312\\
P + W               & 0.931 & 0.930 & 0.928 & 0.927 & 0.333 & 0.334 & 0.327 & 0.313\\
P + W + PCA         & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344\\
Averaging           & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388\\
%Averaging + Directivity & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389\\
\end{tabular}
\caption{TODO \textcolor{red}{version simple!!}}
\label{tab:tab3}
\end{table}

\begin{table}
\small
\centering
\begin{tabular}{@{}l *{8}{c}@{}}
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline \hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
\hline \hline
% Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
Correlation            & 0.886 & 0.884 & 0.891 & 0.877 & 0.153 & 0.145 & 0.170 & 0.132 \\
GENIE3 (depth = 1)     & 0.878 & & & & 0.184 & & & \\
GENIE3 (depth = 2)     & 0.889 & & & & 0.217 & & & \\
GENIE3 (depth = 3)     & & & & & & & & \\
GTE                    & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142 \\
Partial correlation    & 0.930 &  0.927 &  0.926 &  0.923& 0.363  & 0.347 &  0.346 & 0.328 \\
Partial correlation (avg.) & 0.938 & 0.936 & 0.936 & 0.932& 0.391 & 0.389 & 0.385 & 0.374\\
% Our approach simple + dir & 0.939 & 0.936 & 0.937 & 0.933& 0.392 & 0.390 & 0.386 & 0.375\\
Partial correlation (opt.) & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344 \\
Our approach (opt. \& avg.)    & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388 \\
% Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
\end{tabular}
\caption{TODO}
\label{tab:tab1}
\end{table}

@@

@@

In this paper, we present the winning method of the Connectomics challenge. This approach is simple and theorically grounded. We also evaluate the method in a more detailed way and perform a comparison with other inference methods. We also discuss the relevance of the metric used in the challenge.

Obviously, we give a hint about limiting the depth of trees in the GENIE3 method to infer the network. It should be interesting to see if there is a optimum value for the maxmimal depth.

In previous study \citep{shipley2002cause}, it has been suggested to
consider correlation coefficients of all possible orders,
i.e. conditioned on every subsets of the set of $p-2$ other variables. This
would take into account multiple indirect paths between a given pair of
variables. Instead of considering only the highest order, it may be useful to study with accuracy the infered network at all orders. In this case, it may not be very relevant but interesting in case of dymanical networks.

As we mentionned, in order to maximize the ROC AUC, it was not necessary to discover causal relationships. It may be, in future works, promising to focus on that particular topic.
